---
title: "FCR Spatial Sampling 06 Jul 2023"
output: html_notebook
created: "28 Feb 2024"
edited: "1 Mar 2024"

---

On July 6, 2023, we were collecting bathymetery data with the ADCP and deployed an EXO2 off the side of the boat at about the surface or just under the surface. EXO2 data were collected every second. We also collected time and GPS data which we merge together to create the data frame. This data frame does not get saved (if you want it saved add that line). 

After the data have been compiled, you can create some maps using leaflet to look at spatial trends. 


```{r set up}

# Read in packages. Might need to install some. 

pacman::p_load(tidyverse, sf, marmap, leaflet, raster, rgdal)


```

```{r Read in EXO file}

# Read in the EXO2 file 

# Create a vecotr of the names of the variables

EXO_names <- c("Date", "Time", "Sec" ,"Site", "Chlorophyll_RFU", "Cond_uScm", "Depth_m", "fDOM_QSU", "fDOM_RFU", "nLFCond_uScm",
 "ODO_sat","local_ODO_sat" ,"ODO_mgL", "Pressure_psi_g", "Sal_psu", "SpCond_uScm", "PC_RFU", "TDS_mgL", "Turbidity_FNU", "TSS_mgL", "Wiper_Position_V",
 "Temp_C", "Vertical_Position_m"," Battery_V", "Cable_Pwr_V")

# Read in file and eliminate all the header information

EXO <- read.csv("EXO_FCR_spatial_2023_07_06.csv", skip=19, header = F, skipNul = TRUE)

# Add the column names we defined above
colnames(EXO) <- EXO_names

# make a datetime column

EXO$Date <- as.Date(EXO$Date, format="%m/%d/%Y")

EXO$DateTime <- ymd_hms(paste0(EXO$Date," ",EXO$Time))

# make ugL from RFU for Chla and Phyco. Relationship from CCR sensor. 

EXO2 <- EXO%>%
  mutate(
    Chlorophyll_ugL = (Chlorophyll_RFU*4)-0.63,
    PC_ugL = (PC_RFU*1)-0.59
  )%>%
  select(DateTime, everything()) # put datetime column first
```

```{r Read in Coords}

# Read in the Coordinates file from the GPS

Cord <- read_csv("FCR_6JUL2023_TimeAndLocation.csv")

# round to the whole second
Cord$sec2 <- round(Cord$Second, digits = 0)

# Combine the Date and time columns into one column
Cord$DateTime <-ymd_hms(paste0(Cord$Year,"-" ,0,Cord$Month ,"-",0,Cord$Day," ",Cord$Hour, ":",Cord$Minute, ":", Cord$sec2))

# Take the average of the coordinates over the second. We can decide if that is best or pick the coordinate that is closest to the second

Cords2 <- Cord%>%
  group_by(DateTime)%>%
  summarise(across(c(Easting_WGS84_m, Northing_WGS84_m), mean))%>%
  ungroup()

```

```{r group}

# Merge the data frame by DateTime. There is alot more coordinates than EXO2 observations
all_df <- merge(EXO2, Cords2, by="DateTime")

# take out observations when the EXO was out of the water. We define this as less than 40 uScm specific conductivity. 
# Rename the column headers to long and lat
adcpdata <- all_df%>%
  filter(SpCond_uScm>40)%>%
  rename(long = Easting_WGS84_m,
         lat = Northing_WGS84_m)

```

This code is from the ADPC mapping script by Michael K. 


```{r convert coordinate}

# convert coordinates from WGS84 UTM Zone 17 to WGS84 long/lat

# extract UTM coordinates
adcpdata_coords <- adcpdata[c("long", "lat")]

# assign the coordinates to UTM, WGS84, and zone 17, measured in units of meters
adcpdata_coords_utm <- SpatialPoints(adcpdata_coords, proj4string=CRS("+proj=utm +zone=17 +datum=WGS84 +units=m"))
# convert the coordinates to Long/Lat in WGS84
adcpdata_coords_geo <- spTransform(adcpdata_coords_utm, CRS("+proj=longlat +datum=WGS84"))

# re-insert converted coordinates
adcpdata_coords <- coordinates(adcpdata_coords_geo)
adcpdata_coords <- as.data.frame(adcpdata_coords)
adcpdata$long <- adcpdata_coords$coords.x1
adcpdata$lat <- adcpdata_coords$coords.x2

```

```{r select the variable you want to map}


DO <- adcpdata%>%
  dplyr::select(long,lat,Chlorophyll_ugL)%>% # select the variable here 
  filter(Chlorophyll_ugL<50) # might need to filter so very high or very low values
  


```

```{r plot the data}
  
reg <- griddify(DO, nlon = 100, nlat = 300) #nlon and nlat set the number of pixels in the map; play around with these values. 
class(reg) #Returns the class of data

bat <- as.bathy(reg) #Classifies the data as "bathy"
class(bat) #Returns the class of data


bat_map <- plot(bat, image = TRUE, lwd = 0.1) #Creates bathymetric map
bat_map 

b <- marmap::as.raster(bat)
pal <- colorNumeric(c("red", "orange", "#FFFFCC", "green",'aquamarine', "blue", "purple"), # change the number of colors you want
                    domain = DO$Chlorophyll_ugL, # change column name here
                    na.color = "transparent")

# make the map
map <- leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(-79.8375,37.3058, zoom = 16) %>%
  addRasterImage(b, colors = pal) %>%
  addLegend("bottomright", pal = pal, values = DO$Chlorophyll_ugL, title = "Chla (ugL)") # change the column name here
map

# Save the map if you want to 

ggsave("plot.jpg", width = 7, height = 5, units = "in")

```

